Formalization, Mechanization and Automation of \\ G\"{o}del's Proof of God's Existence

### 2013-10-29 Talks/FU-Berlin/GödelGod-Talk.tex
FU talk

\title[Formalizations of G\"{o}del's Proof of God's Existence]{Formalizations of G\"{o}del's Proof of God's Existence}

\author{Christoph Benzm\"{u}ller and Bruno Woltzenlogel Paleo}

\institute[] {
  \inst{}%
}

\date[29.10.13]
{2013 October ??}

@@

\begin{frame}
  \titlepage
\end{frame}

>>>>

\title[On G\"{o}del's Proof of God's
Existence]{Formalization, Mechanization and Automation of \\ G\"{o}del's
  Proof of God's Existence}

\author{\textbf{Christoph Benzm\"{u}ller} and \textbf{Bruno Woltzenlogel Paleo}}

% \institute[]{
%   \inst{}%
% }

\date[1.11.2013]{November 1, 2013}

@@

\begin{frame}
  \titlepage
\includegraphics[height=2.5cm]{buch7.jpg} 
\hfill
\includegraphics[height=2.5cm]{nd.jpg}
\end{frame}

\input{talk-part-1.tex}

\input{talk-part-2.tex}

\input{talk-part-3.tex}

\input{talk-part-4.tex}

\input{talk-part-5.tex}

\input{talk-part-6.tex}

### 2013-10-29 Talks/FU-Berlin/talk-part-1.tex

\ldots\rotatebox[origin = bl,width = 0mm]{65}{\textcolor{blue}{Anselm v. C.}} \hskip-2.3em
\ldots  \rotatebox[origin = bl]{65}{\textcolor{red}{Th. Aquinas}}  \hskip-2.3em
\ldots\ldots  \rotatebox[origin = bl]{65}{\textcolor{blue}{Spinoza}} \hskip-1.3em
                \rotatebox[origin = bl]{65}{\textcolor{blue}{Leibniz}}  \hskip-1.2em
\ldots  \rotatebox[origin = bl]{65}{\textcolor{red}{Hume}}  \hskip-1em
          \rotatebox[origin = bl]{65}{\textcolor{red}{Kant}}  \hskip-.8em
\ldots  \rotatebox[origin = bl]{65}{\textcolor{blue}{Hegel}}  \hskip-1.3em
\ldots  \rotatebox[origin = bl]{65}{\textcolor{red}{Frege}}  \hskip-1.3em
\ldots  \rotatebox[origin = bl]{65}{\textcolor{blue}{Hartshorne}} \hskip-1.9em
          \rotatebox[origin = bl]{65}{\textcolor{blue}{Malcolm}}  \hskip-1.4em
          \rotatebox[origin = bl]{65}{\textcolor{red}{Lewis}}  \hskip-1em
          \rotatebox[origin = bl]{65}{\textcolor{blue}{Plantinga}}  \hskip-1.6em
          \rotatebox[origin = bl]{65}{\textcolor{blue}{G\"odel}}   \hskip-1.2em
\ldots \\[1em]

@@

\begin{frame}{Introduction} \Large
Different Interests in Ontological Arguments: \\[1em]
\begin{itemize}
\item \textcolor{blue}{Philosophical:} Boundaries of Mataphysics \& Epistemology
  \begin{itemize}
  \item We talk about a metaphysical concept (God), 
  \item but we want to draw
      a conclusion for the real world. \\[2em]
  \end{itemize} 
\item \textcolor{blue}{Theistic:} Successful argument should convince atheists. \\[2em]
\item \textcolor{red}{Our:} Can computers (theorem provers) be used
  \begin{itemize}
  \item to formalize the definitions and axioms?
  \item to verify the arguments step-by-step?
  \item to fully automate (sub-)arguments? \\[1em]
  \end{itemize}
  \textcolor{red}{\emph{``Computer-assisted Theoretical Philosophy''}}
\end{itemize}
\end{frame}

\begin{frame}{Introduction} \large
Main Challenge: \hfill No theorem provers for \emph{Second-order Modal
  Logic} \\[1em]

Our Idea: \hfill Exploit an embedding in \emph{Classical Higher-order
  Logic} \\
\,\hfill {\small [Benzm\"ullerPaulson, Logica Universalis, 2013]} \\[2em]

@@

\begin{itemize}
\item[A:] pen and paper: detailed natural deduction proof 
\item[B:] formalization: axioms, defs, thms  in \textsc{TPTP THF}
\item[C:] consistency: automatic verification with \textsc{Nitpick} 
\item[D:] proof automation: theorems provers Leo-II and \textsc{Satallax} 
\item[E:] step-by-step verification: proof assistant \textsc{Coq} 
\item[F:] automation \& verification: proof assistant
  \textsc{Isabelle}
\item[ ] Conclusion 
\end{itemize}
Did we get anything new? \ldots Probably yes --- let's discuss afterwards!
\end{frame}

>>>>

\hskip-.5em
\ldots\rotatebox[origin = bl,width = 0mm]{65}{\textcolor{blue}{Anselm v. C.}} \hskip-2.3em
\ldots  \rotatebox[origin = bl]{65}{\textcolor{red}{Th. Aquinas}}  \hskip-2.3em
\ldots\ldots   \rotatebox[origin = bl]{65}{\textcolor{blue}{Descartes}} \hskip-1.7em
               \rotatebox[origin = bl]{65}{\textcolor{blue}{Spinoza}} \hskip-1.3em
               \rotatebox[origin = bl]{65}{\textcolor{blue}{Leibniz}}  \hskip-1.2em
\ldots  \rotatebox[origin = bl]{65}{\textcolor{red}{Hume}}  \hskip-1em
          \rotatebox[origin = bl]{65}{\textcolor{red}{Kant}}  \hskip-.8em
\ldots  \rotatebox[origin = bl]{65}{\textcolor{blue}{Hegel}}  \hskip-1.3em
\ldots  \rotatebox[origin = bl]{65}{\textcolor{red}{Frege}}  \hskip-1.3em
\ldots  \rotatebox[origin = bl]{65}{\textcolor{blue}{Hartshorne}} \hskip-1.9em
          \rotatebox[origin = bl]{65}{\textcolor{blue}{Malcolm}}  \hskip-1.4em
          \rotatebox[origin = bl]{65}{\textcolor{red}{Lewis}}  \hskip-1em
          \rotatebox[origin = bl]{65}{\textcolor{blue}{Plantinga}}  \hskip-1.6em
          \rotatebox[origin = bl]{65}{\textcolor{blue}{G\"odel}}   \hskip-1.2em
\ldots \\[1em]

@@

\begin{frame}{Introduction} \Large
Different Interests in Ontological Arguments: \\[1em]
\begin{itemize}
\item \textcolor{blue}{Philosophical:} Boundaries of Metaphysics \& Epistemology
  \begin{itemize}
  \item We talk about a metaphysical concept (God), 
  \item but we want to draw
      a conclusion for the real world. \\[1em]
  \item Necessary Existence (NE): metaphysical NE vs. logical NE  vs. modal NE \\[2em]
  \end{itemize} 
\item \textcolor{blue}{Theistic:} Successful argument should convince atheists. \\[2em]
\item \textcolor{red}{Our:} Can computers (theorem provers) be used
  \begin{itemize}
  \item to formalize the definitions and axioms?
  \item to verify the arguments step-by-step?
  \item to fully automate (sub-)arguments? \\[1em]
  \end{itemize}
  \textcolor{red}{\emph{``Computer-assisted Theoretical Philosophy''}}
\end{itemize}
\end{frame}

\begin{frame}{Introduction} \large
Main challenge: \hfill No theorem provers for \emph{Higher-order Modal Logic} \\[1em]

Our idea: \hfill Exploit an embedding in \emph{Higher-order Classical Logic} \\
\,\hfill {\small [Benzm\"ullerPaulson, Logica Universalis, 2013]} \\[2em]

@@

\begin{itemize}
\item[A:] pen and paper: detailed natural deduction proof 
\item[B:] formalization: axioms, defs, thms  in \textsc{TPTP THF}
\item[C:] consistency: automatic verification with \textsc{Nitpick} 
\item[D:] proof automation: theorems provers Leo-II and \textsc{Satallax} 
\item[E:] step-by-step verification: proof assistant \textsc{Coq} 
\item[F:] automation \& verification: proof assistant 
  \textsc{Isabelle} \\[2em]
%\item[ ] Conclusion \\[2em]
\end{itemize}
Did we get new results? \hfill  \textcolor{red}{Yes --- let's discuss later!}
\end{frame}
